### 2023-06-28 analysis/epi_analysis.Rmd
update results

---
title: "Epidemiological Data Analysis"
author: "Nicolas Banholzer"
date: "2023-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(tidybayes)
library(lubridate)
library(rstan)
source("../utils/epi.r")
source("../utils/bayes.r")
source("../utils/plotting.r")
source("../utils/tex.r")
```

## Data

```{r}
epi_data <- readRDS("../data-clean/epidemiological.rds")
epi_data_ll <- readRDS("../data-clean/redcap-long.rds")
env_data <- readRDS("../data-clean/environmental.rds")
mol_data <- readRDS("../data-clean/molecular.rds") 
sec_data <- readRDS("../data-clean/secondary.rds")
```

## Descriptives

### Baseline data

```{r}
# absent person days
absent_pers_days <- epi_data_ll %>%
  mutate(absent_days = as.numeric(date_back - date_absent),
         reason = ifelse(as.character(reason) == "unknown", "other", as.character(reason))) %>%
  group_by(class, reason) %>%
  summarize(n = sum(absent_days)) %>%
  ungroup() %>%
  group_by(class) %>%
  mutate(p = n / sum(n)) %>%
  ungroup()
absent_pers_days <- absent_pers_days %>%
  add_row(class = "A", reason = "total", n = sum(absent_pers_days$n[absent_pers_days$class=="A"]),
          p = sum(absent_pers_days$n[absent_pers_days$class=="A"]) / sum(absent_pers_days$n)) %>%
  add_row(class = "B", reason = "total", n = sum(absent_pers_days$n[absent_pers_days$class=="B"]),
          p = sum(absent_pers_days$n[absent_pers_days$class=="B"]) / sum(absent_pers_days$n)) 
absent_pers_days <- absent_pers_days %>%
  add_row(class = "Total", reason = "total", n = sum(absent_pers_days$n[absent_pers_days$reason=="total"]),
          p = 1) %>%
  add_row(class = "Total", reason = "sickness", n = sum(absent_pers_days$n[absent_pers_days$reason=="sickness"]),
          p = sum(absent_pers_days$n[absent_pers_days$reason=="sickness"])  / sum(absent_pers_days$n[absent_pers_days$reason=="total"])) %>%
  add_row(class = "Total", reason = "other", n = sum(absent_pers_days$n[absent_pers_days$reason=="other"]),
          p = sum(absent_pers_days$n[absent_pers_days$reason=="other"])  / sum(absent_pers_days$n[absent_pers_days$reason=="total"])) %>%
  mutate(p = round(100 * p)) %>%
  mutate(class = factor(class, levels = c("A", "B", "Total")),
         reason = factor(reason, levels = c("total", "sickness", "other"))) %>%
  arrange(class, reason) 
absent_pers_days

# total absent
round(absent_pers_days$n[absent_pers_days$class=="Total"] / (7 * 5 * 20 + 7 * 5 * 18) * 100)

# epidemiological cases
epi_data %>%
  group_by(class) %>%
  summarize(n = sum(new_cases)) %>%
  ungroup() %>%
  mutate(p = n / sum(n) * 100 %>% round) %>%
  add_row(class = "Total", n = sum(.$n), p = 100)

# molecular cases
mol_data %>%
  mutate(Pos = N - Neg) %>%
  group_by(class) %>%
  summarize(n = sum(Pos)) %>%
  ungroup() %>%
  mutate(p = n / sum(n) * 100 %>% round) %>%
  add_row(class = "Total", n = sum(.$n), p = 100)
```

### Line List

```{r}
line_list_dat <- epi_data_ll %>%
  filter(respiratory_infection == "yes") %>%
  dplyr::select(class, date_symptom, date_absent, date_back) %>%
  arrange(class, date_absent) %>%
  as.matrix() %>%
  xtable::xtable()

print.xtable(line_list_dat,
             type = "latex",
             file = "../results/epi-data/line-list-data.tex",
             sanitize.text.function = identity,
             include.colnames = F,
             include.rownames = F,
             only.contents = T,
             hline.after = NULL)
```

### Number of cases

```{r}
# number of cases by class
epi_data %>%
  filter(vacation == 0) %>%
  group_by(class) %>%
  summarize(cases = sum(new_cases))

# number of cases by intervention
epi_data %>%
  filter(vacation == 0) %>%
  group_by(aircleaner) %>%
  summarize(cases = sum(new_cases))

# number of cases by week
cases_by_week_pl <- epi_data %>%
  mutate(week = isoweek(date)) %>%
  filter(week != 6) %>%
  mutate(week = ifelse(week < 6, week - min(week) + 1, week - min(week))) %>%
  group_by(class, week) %>%
  summarize(cases = sum(new_cases),
            aircleaner = first(aircleaner)) %>%
  ungroup() %>%
  mutate(aircleaner = factor(aircleaner, levels = c("Yes", "No")),
         class = factor(class, levels = c("A", "B"))) %>%
  arrange(class, aircleaner) %>%
  ggplot(aes(x = factor(week), y = cases, fill = aircleaner, pattern = class)) +
  geom_bar_pattern(stat = "identity", color = "black", pattern_fill = "white", pattern_angle = 45,
                   pattern_density = 0.25, pattern_spacing = 0.05, width = .5) +
  labs(x = "Study week", y = "Number of cases", fill = "Air cleaner", pattern = "Class", title = "a") +
  scale_fill_manual(values = rev(wes_palette("Royal1")[1:2])) +
  scale_x_discrete(breaks = 1:7) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = seq(0, 10, 2)) +
  scale_pattern_manual(values = c(A = "stripe", B = "none")) +
  theme_custom() +
  theme(legend.position = c(0.13, 0.8),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.3, "cm"),
        plot.title.position = "plot") +
  guides(fill = guide_legend(nrow = 1, override.aes = list(pattern = "none", order = 2)),
         pattern = guide_legend(nrow = 1, override.aes = list(fill = "white"), order = 1))

save_plot(cases_by_week_pl,
          pdf_file = "../results/epi-data/cases_by_week.pdf",
          eps_file = "../results/epi-data/cases_by_week.eps",
          w = 16, h = 8) 
```

## Prepare data

### Assumptions

* Incubation periods are informed by https://www.sciencedirect.com/science/article/pii/S1473309909700696?casa_token=MSYyhmCLcLAAAAAA:10ooZqTNmMMfk6sIB1idSETWMusOs3Y-tKWPqGUXOFiD0PPFB0edT3W5OZyDO13S6kWPTv8NDOU and for SARS-CoV-2 from https://bmjopen.bmj.com/content/bmjopen/10/8/e039652.full.pdf.
* We assume a lognormal distribution and use the median and dispersion estimate
* We weight the incubation period distributions each week according to the proportion of positive samples for each virus --> effective incubation period.
* For Metapneumovirus, there was insufficient data. Here we take the mean of the parameters of the other viruses.

```{r}
# mean estimate
p_in_mean_m <- c(
  5.6, # AdV
  0.6, # Flu B
  2.6, # PIV
  1.9 # HRV
)

# 95% limits of estimate
p_in_mean_ul <- matrix(c(
  6.3, 4.8,
  0.6, 0.5,
  3.1, 2.1,
  2.4, 1.4
), ncol = 2, byrow = T)

# mean dispersion
p_in_sd_m <- c(
  1.26,
  1.51,
  1.35,
  1.68
)

# upper and lower limit of SD
p_in_sd_ul <- matrix(c(
  1.38, 1.13,
  1.64, 1.37,
  1.55, 1.16,
  2.01, 1.16
), ncol = 2, byrow = T)

# Convert to lognormal distr parameters
p_in_mu_m <- c(log(p_in_mean_m), 1.63) # the last value is for SARS-CoV-2
p_in_mu_s <- c(log(p_in_mean_ul[,1] / p_in_mean_ul[,2]), (1.75 - 1.51)) / (2 * qnorm(.975))
p_in_sigma_m <- c(log(p_in_sd_m), 0.5)
p_in_sigma_s <- c(log(p_in_sd_ul[,1] / p_in_sd_ul[,2]), (0.55 - 0.46)) / (2 * qnorm(.975))

# visualize
set.seed(12345)

prior_pin_par <- data.frame(
  virus = c("AdV", "Flu B", "PIV", "HRV", "CoV", "MPV"),
  mum = c(p_in_mu_m, mean(p_in_mu_m)), 
  mus = c(p_in_mu_s, mean(p_in_mu_s)),
  sim = c(p_in_sigma_m, mean(p_in_sigma_m)),
  sis = c(p_in_sigma_s, mean(p_in_sigma_s))
) %>%
  mutate(virus = factor(virus, levels = c("Flu B", "HRV", "AdV", "CoV", "MPV", "PIV"))) %>%
  arrange(virus)

prior_pin <- prior_pin_par %>%
  mutate(mu = map2(mum, mus, function(m, s) rnorm(n = 4e3, mean = m, sd = s)),
         sigma = map2(sim, sis, function(m, s) rtrunc(n = 4e3, spec = "norm", a = 0, b = Inf, mean = m, sd = s))) %>%
  unnest() %>%
  mutate(pin = map2(mu, sigma, function(log_mu, log_sigma) data.frame(x = 0:10, y = vp(10, T, p_in, log_mu, log_sigma)))) %>%
  unnest()

prior_pin_pl <- prior_pin %>%
  ggplot(aes(x = x, y = y)) +
  stat_lineribbon(point_interval = "mean_qi") +
  scale_fill_brewer() + 
  facet_wrap(~ virus) +
  labs(x = "s (delay in days)", y = expression(p[IN]*" (%)")) +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 10, 2)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  theme_bw2() +
  theme(plot.title = element_text(hjust = 0), legend.position = "none")
prior_pin_pl

save_plot(prior_pin_pl, pdf_file = "../results/epi-data/incubation-periods.pdf", w = 16, h = 10)
```

### Stan data

```{r}
# stan data
stan_data <- list()

# number of seeding days
stan_data$S <- ceiling(2 * exp(max(p_in_mu_m) + max(p_in_sigma_m)^2/2))

# number of days from Saturday to study start
stan_data$SS <- seq(min(epi_data$date) - days(7), min(epi_data$date), by = "1 day")
stan_data$SS <- abs(as.numeric(stan_data$SS[weekdays(stan_data$SS) == "Saturday"] - min(min(epi_data$date))))

# number of new cases
stan_data$cases <- epi_data %>%
  dplyr::select(class, date, new_cases) %>%
  dcast(date ~ class) %>%
  arrange(date) %>%
  dplyr::select(A, B) %>%
  as.matrix()

# number of classes
stan_data$C <- ncol(stan_data$cases)

# number of days and weeks
stan_data$D <- nrow(stan_data$cases)
stan_data$W <- ceiling(stan_data$D / 7)

# total cases by weekday
stan_data$totCasesWeekday <- epi_data %>%
  group_by(weekday) %>%
  summarize(n = sum(new_cases)) %>%
  ungroup() %>%
  mutate(weekday = factor(weekday, levels = c("Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))) %>%
  arrange(weekday) %>%
  dplyr::select(n) %>%
  unlist()

# weekend
stan_data$weekend <- epi_data %>%
  filter(class == "A") %>%
  dplyr::select(weekend) %>%
  unlist()
seeding_weekend <- epi_data$date[epi_data$class=="A"][1] - days(stan_data$S:1)
seeding_weekend <- ifelse(weekdays(seeding_weekend) %in% c("Saturday", "Sunday"), 1, 0)
stan_data$weekend <- c(seeding_weekend, stan_data$weekend)

# vacation
stan_data$vacation <- epi_data %>%
  filter(class == "A") %>%
  dplyr::select(vacation) %>%
  unlist()
stan_data$vacation <- c(rep(0, stan_data$S), stan_data$vacation)

# aircleaner intervention
stan_data$airCleaner <- epi_data %>%
  mutate(aircleaner = ifelse(aircleaner == "Yes", 1, 0)) %>%
  dplyr::select(class, date, aircleaner) %>%
  dcast(date ~ class) %>%
  arrange(date) %>%
  dplyr::select(A, B) %>%
  as.matrix()
seeding_airCleaner <- matrix(0, nrow = stan_data$S, ncol = stan_data$C) 
stan_data$airCleaner <- rbind(seeding_airCleaner, stan_data$airCleaner)

# proportion of students in class
stan_data$stud <- epi_data %>%
  mutate(prop_in_class = n_present / n_class,
         prop_in_class = (prop_in_class - mean(prop_in_class)) / sd(prop_in_class)) %>%
  dplyr::select(class, date, prop_in_class) %>%
  dcast(date ~ class) %>%
  arrange(date) %>%
  dplyr::select(A, B) %>%
  as.matrix()

# air change rate
stan_data$vent <- epi_data %>%
  dplyr::select(class, date, vacation, weekend) %>%
  left_join(env_data %>% dplyr::select(class, date, AER), by = c("class", "date")) %>%
  mutate(AER = (AER - mean(AER, na.rm = T)) / sd(AER, na.rm = T)) %>%
  mutate(AER = ifelse(vacation + weekend > 0, -1001, AER),
         AER = ifelse(is.na(AER), +1001, AER)) %>%
  dplyr::select(class, date, AER) %>%
  dcast(date ~ class) %>%
  arrange(date) %>%
  dplyr::select(A, B) %>%
  as.matrix()
stan_data$missingVent <- apply(stan_data$vent, c(1,2), function(x) ifelse(x == 1001, 1, 0))
stan_data$Mvent <- sum(stan_data$missingVent)
stan_data$missingVent <- apply(stan_data$missingVent, 2, function(x) {
  y <- cumsum(x)
  dy <- c(0, diff(y))
  y <- ifelse(dy == 0, 0, y)
  return(y)
})

# share of positive covid-19 tests in the community
stan_data$cov <- sec_data %>%
  filter(between(date, min(epi_data$date - days(stan_data$S)), max(epi_data$date))) %>%
  mutate(covid_prop_pos = (covid_prop_pos - mean(covid_prop_pos)) / sd(covid_prop_pos)) %>%
  dplyr::select(covid_prop_pos) %>%
  unlist()

# consultations for influenza-like in the community
stan_data$ili <- sec_data %>%
  filter(between(date, min(epi_data$date - days(stan_data$S)), max(epi_data$date))) %>%
  mutate(ili_consult = (ili_consult - mean(ili_consult)) / sd(ili_consult)) %>%
  dplyr::select(ili_consult) %>%
  unlist()

# number of viruses
stan_data$K <- 6

# positive molecular samples by week and class 
mol_class_week <- mol_data %>%
  mutate(week = week(date),
         week = week - min(week) + 1,
         week = ifelse(week > 4, week - 1, week))  %>%
  dplyr::select(week, class, AdV, `Flu B`, HRV, MPV, PIV, `SARS-CoV-2`) %>%
  melt(c("week", "class")) %>%
  group_by(class, week, variable) %>%
  summarize(n = sum(value)) %>%
  ungroup() %>%
  mutate(variable = as.character(variable), 
         variable = ifelse(variable == "SARS-CoV-2", "CoV", variable),
         variable = factor(variable, levels = c("Flu B", "HRV", "AdV", "CoV", "MPV", "PIV")))
mol_class_week_extend <- epi_data %>%
  dplyr::select(class, date) %>%
  mutate(week = week(date),
         week = week - min(week) + 1,
         week = ifelse(week > 4, week - 1, week))  %>%
  left_join(mol_class_week) %>%
  arrange(class, date) %>%
  dplyr::select(class, date, variable, n) %>%
  dcast(class + date ~ variable)
stan_data$pos_samples <- simplify2array(by(mol_class_week_extend, mol_class_week_extend$class, as.matrix))
stan_data$pos_samples <- apply(stan_data$pos_samples[,-c(1,2),], c(1,2,3), as.integer)

# hyperparameters for prior for incubation period distributions
stan_data$p_in_mu_m <- prior_pin_par$mum
stan_data$p_in_mu_s <- prior_pin_par$mus
stan_data$p_in_sigma_m <- prior_pin_par$sim
stan_data$p_in_sigma_s <- prior_pin_par$sis
```

## Modeling

```{r}
# stan model code file
stan_file <- "../models/latent-infection-model.stan"

# run model
latInfMod <- stan(stan_file, data = stan_data, seed = 12345, cores = 4)

# quick estimation results
summary(latInfMod, probs = c(0.025, 0.5, 0.975), pars = c("phi", "alpha", "omega", "beta", "vega", "mu_p_in", "sigma_p_in", "I0"))$summary %>%
  data.frame() %>%
  mutate_all(round, 2)
```

## Results

### Adjusted Rate Ratio

```{r}
arr <- spread_draws(latInfMod, `beta[1]`) %>%
  rename(effect = `beta[1]`) %>%
  dplyr::select(effect) %>%
  mean_qi(.width = c(0.5, 0.8, .95)) %>%
  mutate(across(c(effect, .lower, .upper), exp))
arr
```

### Avoided infections

```{r}
types <- c("Air cleaners partially installed",
           "Scenario 2: Air cleaners not installed",
           "Scenario 1: Air cleaners installed throughout the study")

avoid_infections <- spread_draws(latInfMod, mu_cases_ac[day,class,type]) %>%
  group_by(.draw, type) %>%
  summarize(infections = sum(mu_cases_ac)) %>%
  ungroup() %>%
  group_by(type) %>%
  mean_qi(.width = c(0.5, 0.8, 0.95)) %>%
  ungroup() %>%
  mutate(type = ifelse(type == 1, types[1], ifelse(type == 2, types[2], types[3])),
         type = factor(type, levels = types[c(3,1,2)])) %>%
  arrange(desc(.width)) %>%
  mutate(`.width` = factor(`.width`, levels = rev(c(.5, .8, .95)))) %>%
  filter(type != types[1])

avoid_infections_pl <- avoid_infections  %>%
  ggplot() +
  geom_rect(aes(xmin = .lower, xmax = .upper, ymin = 0.75, ymax = 1.25, fill = `.width`)) +
  geom_segment(aes(x = infections, xend = infections, y = 0.75, yend = 1.25), size = 1.5 / cm(1)) +
  geom_point(aes(x = infections, y = 1), size = 3, fill = "white", shape = 21) +
  facet_wrap(~ type, ncol = 1) +
  scale_fill_manual(values = wes_palette("Royal1")[c(3,4,2)]) +
  scale_y_continuous(expand = expansion(mult = c(0.25, 0.25))) +
  scale_x_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0)), breaks = seq(0, 200, 25)) +
  labs(x = "Estimated number of infections", fill = "CrI", title = "b") +
  theme_custom() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(color = "white"),
        axis.ticks.y = element_blank(),
        legend.position = "top", legend.direction = "horizontal",
        legend.key.height = unit(.25, "cm"), legend.key.width = unit(.35, "cm"), 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title.position = "plot",
        panel.spacing = unit(20, units = "pt"),
        strip.text.x = element_text(margin = margin(b = 5)))

save_plot(avoid_infections_pl, 
          pdf_file = "../results/epi-data/avoided-infections.pdf",
          eps_file = "../results/epi-data/avoided_infections.eps",
          w = 16, h = 8)
```

```{r}
barplot_mol <- readRDS("../results/mol-data/descriptives-cond.rds") + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE, nrow = 1)) +
  theme(legend.position = "bottom", legend.key.height = unit(.1, "cm"), legend.key.width = unit(.5, "cm"), plot.title.position = "plot") +
  ggtitle("a | Number of positive saliva samples by study condition.",
          subtitle = "Influenza B (Flu B), rhinovirus (HRV), adenovirus (AdV), SARS-CoV-2 (CoV), metapneumovirus (MPV), parainfluenza (PIV).")
boxplot_env <- readRDS("../results/env-data/particles-boxplot.rds") +
  ggtitle("b | Aerosol number (CN) and particle matter (PM) concentrations by study condition.",
          subtitle = expression("Particle diameter between 175nm and 20"*mu*"m. PM for sizes <1 to <10"*mu*"m."))
avoid_infections_pl <- avoid_infections_pl +
  ggtitle("c | Estimated number of infections with and without air cleaners. ",
          subtitle = "Counterfactual scenarios with air cleaners always (never) installed, based on the model-estimated effect of air cleaners.")
env_epi_comb_pl <- grid.arrange(barplot_mol, 
                                grid.arrange(boxplot_env, avoid_infections_pl, nrow = 2, heights = c(8, 8), widths = 16),
                                nrow = 2, heights = c(6, 16), widths = 16)
ggsave(plot = env_epi_comb_pl, "../doc/Epidemics2023/figure.png", width = 16 / cm(1), height = 22 / cm(1))
```

### Estimation results

```{r}
est_res <- summary(latInfMod, probs = c(0.025, 0.975), pars = c("phi", "alpha", "omega", "beta", "vega", "mu_p_in", "sigma_p_in"))$summary %>%
  data.frame() %>%
  add_rownames() %>%
  rename(variable = rowname) %>%
  mutate(variable = c("$\\phi$",
                      "$\\alpha$",
                      "$\\omega$",
                      "$\\beta_1$",
                      "$\\beta_2$",
                      "$\\beta_3$",
                      "$\\beta_4$",
                      "$\\beta_5$",
                      "$\\beta_6$",
                      "$\\vartheta_{\\text{Saturday}}$",
                      "$\\vartheta_{\\text{Sunday}}$",
                      "$\\vartheta_{\\text{Monday}}$",
                      "$\\vartheta_{\\text{Tuesday}}$",
                      "$\\vartheta_{\\text{Wednesday}}$",
                      "$\\vartheta_{\\text{Thursday}}$",
                      "$\\vartheta_{\\text{Friday}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{Flu B}}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{HRV}}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{AdV}}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{CoV}}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{MPV}}}$",
                      "$\\mu_{p_{\\text{IN}}^{\\text{PIV}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{Flu B}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{HRV}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{AdV}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{CoV}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{MPV}}}$",
                      "$\\sigma_{p_{\\text{IN}}^{\\text{PIV}}}$")) %>%
  dplyr::select(variable, mean, `X2.5.`, `X97.5.`, Rhat, `n_eff`) %>%
  mutate(across(c(mean, `X2.5.`, `X97.5.`, Rhat), function(x) tex_sign(round_k(x))),
         `n_eff` = round(`n_eff`)) %>%
  as.matrix()

print.xtable(xtable(est_res),
             type = "latex",
             file = "../results/epi-data/estimation-results.tex",
             sanitize.text.function = identity,
             include.colnames = F,
             include.rownames = F,
             only.contents = T,
             hline.after = NULL)
```

### Model fit

```{r}
fitted_cases <- spread_draws(latInfMod, muCases[day,class]) %>%
  mutate(class = ifelse(class == 1, "A", "B")) %>%
  filter(day > stan_data$SS) %>%
  mutate(day = day - stan_data$SS) %>%
  left_join(epi_data %>%
              dplyr::select(class, date, new_cases) %>%
              group_by(class) %>%
              arrange(date) %>%
              mutate(day = 1:n()) %>%
              ungroup(),
            by = c("class", "day")) %>%
  mutate(week = isoweek(date)) %>%
  filter(week != 6) %>%
  ungroup() %>%
  mutate(week = ifelse(week < 6, week - min(week) + 1, week - min(week))) %>%
  group_by(week, .draw) %>%
  summarise(estimated_cases = sum(muCases),
            true_cases = sum(new_cases)) %>%
  ungroup()

fitted_cases_pl <- ggplot(mapping = aes(x = week)) +
  stat_lineribbon(data = fitted_cases,
                  mapping = aes(y = estimated_cases),
                  .width = c(0.5, 0.8, 0.95), color = "brown") +
  geom_line(data = fitted_cases %>% group_by(week) %>% slice(1),
            mapping = aes(y = true_cases, color = "Observed"), size = 1.5) +
  scale_fill_manual(values = wes_palette("Royal1")[c(3,4,2)]) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), breaks = scales::breaks_width(2)) +
  scale_color_manual(values = "black") +
  labs(y = "Number of new infections", x = "Study week", fill = "Estimated (CrI)") +
  theme_custom() +
  theme(legend.position = "top") +
  guides(color = guide_legend(title = element_blank()))
fitted_cases_pl

save_plot(fitted_cases_pl, pdf_file = "../results/epi-data/model-fit.pdf",
          w = 16, h = 10)
```