### 2023-10-04 analysis/study_comaprison.Rmd

---
title: "Study Comparison"
author: "Nicolas Banholzer"
date: "2023-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
source("../utils/epi.r")
source("../utils/bayes.r")
source("../utils/plotting.r")
source("../utils/tex.r")
```

## Virus distribution

```{r}
saliva <- data.frame(
    year = c(rep(2022, 6), rep(2023, 6)),
    virus = rep(c("Flu", "HRV", "AdV", "CoV", "MPV", "PIV"), 2),
    n = c(c(1, NA, 1, 19, NA, NA), c(15, 15, 14, 3, 2, 1))
  ) %>%
  group_by(year) %>%
  mutate(p = round(100 * n / sum(n, na.rm = T))) %>%
  ungroup() %>%
  mutate(year = factor(year, levels = c(2022, 2023)),
         virus = factor(virus, levels = c("Flu", "HRV", "AdV", "CoV", "MPV", "PIV"))) %>%
  ggplot(mapping = aes(y = factor(year), x = p, fill = virus)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE), width = .5) +
  geom_text(mapping = aes(label = p), size = 8 / cm(1), color = "white", position = position_stack(vjust = 0.5, reverse = T)) +
  scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
  scale_fill_manual(values = wes_palette("IsleofDogs1")) +
  labs(x = "Proportion of positive saliva samples (%)", y = "Study year", fill = "Virus", title = "a") +
  theme_custom() +
  theme(legend.position = "top",
        plot.title.position = "plot",
        legend.key.height = unit(.1, "cm"),
        legend.key.width = unit(.3, "cm")) +
  guides(fill = guide_legend(nrow = 1))

saliva
```

## No. of positive samples

```{r}
pos_samples <- data.frame(
  year = factor(c(rep(2022, 3), rep(2023, 3)), levels = c(2022, 2023)),
  type = factor(rep(c("Saliva", "Bioaerosols", "HEPA-filter"), times = 2), levels = c("Saliva", "Bioaerosols", "HEPA-filter")),
  n = c(c(21, 10, 8), c(50, 2, 4))
) %>%
  ggplot(aes(x = n, y = year, pattern = year, color = year, pattern_color = year)) +
  geom_bar_pattern(stat = "identity", fill = "white", pattern_density = 0.1, pattern_spacing = 0.1, width = .75) +
  geom_text(aes(label = n), size = 8 / cm(1), hjust = -1) +
  facet_wrap(~ type, ncol = 1) +
  scale_x_continuous(expand = expansion(add = c(0,9.5))) +
  scale_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) +
  scale_pattern_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) +
  labs(y = "Study year", x = "Frequency of positive samples", title = "b") +
  theme_custom() +
  theme(legend.position = "none",
        plot.title.position = "plot")

pos_samples
```

## Particle reduction

```{r}
vars <- c("CO2", "rH", "Temp", "AER", "Cn", "PM1", "PM25", "PM4", "PM10")
vars_labels_no_units <- c("CO[2]", "Relative~humidity", "Temperature", "Air~exchange~rate",
                          "CN", "PM[1]", "PM[2.5]", "PM[4]", "PM[10]")
names(vars_labels_no_units) <- vars

particle <- data.frame(
  year = factor(c(rep(2022, 15), rep(2023, 15)), levels = c(2022, 2023)),
  variable = factor(c(rep(c("Cn", "PM1", "PM25", "PM4", "PM10"), each = 3, times = 2)), levels = c("Cn", "PM1", "PM25", "PM4", "PM10")),
  interval = rep(c("Mean", "Lower", "Upper"), times = 5),
  estimate = c(c(39.24, 3.88, 68.80, 
                 41.05, 6.62, 67.98,
                 45.12, 19.66, 63.16,
                 46.49, 27.26, 61.72,
                 46.60, 29.99, 59.49),
               c(76.83, 63.20, 86.28,
                 76.79, 64.84, 85.46,
                 72.86, 64.24, 79.66,
                 70.77, 63.46, 76.97,
                 99.99, 99.94, 100.00))
) %>%
  dcast(year + variable ~ interval) %>%
  ggplot(aes(y = year, color = year)) +
  facet_wrap(~ variable, nrow = 1, dir = "h", labeller = as_labeller(vars_labels_no_units, label_parsed)) +
  geom_errorbar(mapping = aes(xmin = Lower, xmax = Upper), lineend = "square", size = 1, width = .05) +
  geom_point(mapping = aes(x = Mean), shape = 23, size = 2, fill = "white") +
  scale_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), breaks = c(0, 50, 100), limits = c(0, 100)) +
  labs(y = "Study year", x = "Estimated reduction in particle concentration (%)", title = "c") + 
  theme_custom() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        panel.spacing = unit(.5, "lines"))

particle
```

## Risk ratio

```{r}
rr <- data.frame(
  year = factor(c(rep(2022, 3), rep(2023, 3)), levels = c(2022, 2023)),
  interval = rep(c("Mean", "Lower", "Upper"), times = 2),
  estimate = c(c(1.00, 0.22, 2.067), c(0.72, 0.44, 1.15))
) %>%
  dcast(year ~ interval) %>%
  ggplot(aes(y = year, color = year)) +
  geom_vline(aes(xintercept = 1), linetype = "dashed") +
  geom_errorbar(mapping = aes(xmin = Lower, xmax = Upper), lineend = "square", size = 1, width = .05) +
  geom_point(mapping = aes(x = Mean), shape = 23, size = 2, fill = "white") +
  labs(y = "Study year", x = "ARR of infection with air cleaners (%)", title = "d") + 
  scale_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) +
  theme_custom() +
  theme(legend.position = "none",
        plot.title.position = "plot")

rr  
```

## Combined

```{r}
comparison <- grid.arrange(saliva, pos_samples, particle, rr, ncol = 2, nrow = 2, heights = c(6, 4), widths = c(10,6))
ggsave(plot = comparison, file = "../results/comparison.pdf", width = 16 / cm(1), height = 10 / cm(1))
```