### 2023-10-16 analysis/env_analysis.Rmd

---
title: "Environmental Data Analysis"
author: "Nicolas Banholzer"
date: "2023-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(rstanarm)
library(tidybayes)
source("../utils/plotting.r")
source("../utils/tex.r")
```

## Data

```{r}
env_data <- readRDS("../data-clean/environmental.rds")
epi_data <- readRDS("../data-clean/epidemiological.rds")

env_data <- na.omit(env_data)
env_epi_data <- left_join(
  env_data,
  epi_data %>% 
    group_by(class) %>%
    arrange(date) %>%
    mutate(cases = cumsum(new_cases)) %>%
    dplyr::select(class, date, weekday, n_present, cases),
  by = c("class", "date")
)

vars <- c("CO2", "rH", "Temp", "AER", "Cn", "PM1", "PM25", "PM4", "PM10")
vars_labels <- c("CO[2]~(ppm)", "Relative~humidity~('%')", "Temperature~(degree*C)", 
                 "AER~(air~changes/h)", "CN~(1/cm^3)", "PM[1]~(mu*gm^{-3})", "PM[2.5]~(mu*gm^{-3})", "PM[4]~(mu*gm^{-3})", "PM[10]~(mu*gm^{-3})")
names(vars_labels) <- vars
vars_labels_no_units <- c("CO[2]", "Relative~humidity", "Temperature", "Air~exchange~rate",
                          "CN", "PM[1]", "PM[2.5]", "PM[4]", "PM[10]")
names(vars_labels_no_units) <- vars
```

## Summaries

```{r}
# overall
env_epi_data %>%
  summarize(across(c(CO2, AER, Cn, PM1), .fns = list(mean = mean, sd = sd), .names = "{.col}_{.fn}")) %>%
  mutate(across(c(CO2_mean, CO2_sd, Cn_mean, Cn_sd), round),
         across(c(AER_mean, AER_sd, PM1_mean, PM1_sd), round, 2))

tt_co2 <- t.test(env_epi_data$CO2[env_epi_data$aircleaner=="Yes"], env_epi_data$CO2[env_epi_data$aircleaner=="No"])

# by air cleaner
env_epi_data %>%
  group_by(aircleaner) %>%
  summarize(across(c(CO2, AER, Cn, PM1), .fns = list(mean = mean, sd = sd), .names = "{.col}_{.fn}")) %>%
  ungroup() %>%
  mutate(across(c(CO2_mean, CO2_sd, Cn_mean, Cn_sd), round),
         across(c(AER_mean, AER_sd, PM1_mean, PM1_sd), round, 2))
```

## Boxplots

```{r}
# only particle number concentration
boxpl_cn <- env_epi_data %>%
  dplyr::select(class, date, aircleaner, Cn) %>%
  ggplot(aes(x = aircleaner, fill = aircleaner, y = Cn)) +
  geom_boxplot2() +
  #geom_jitter(alpha = .3, width = .2, color = "grey") +
  scale_fill_manual(values = wes_palette("Royal1")) +
  scale_shape_manual(values = c(2,20)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA)) +
  labs(title = "b", y = expression("Aerosol number concentration (1/cm"^3*")"), x = "Air cleaner") +
  theme_custom() +
  theme(legend.position = "none",
        plot.title.position = "plot") +
  guides(color = "none")

boxpl_cn

save_plot(boxpl_cn, 
          pdf_file = "../results/env-data/aerosol-number-boxplot.pdf", 
          eps_file = "../results/env-data/aerosol-number-boxplot.eps", 
          w = 8, h = 8)
```

```{r}
# particle concentrations
boxpl_pc <- env_epi_data %>%
  dplyr::select(class, date, weekday, aircleaner, Cn, PM1, PM25, PM4, PM10) %>%
  melt(c("class", "date", "weekday", "aircleaner")) %>%
  rename(`Air Cleaner` = aircleaner,
         Class = class) %>%
  ggplot(aes(x = `Air Cleaner`, color = `Air Cleaner`, y = value)) +
  facet_wrap(~ variable, scales = "free_y", nrow = 1, dir = "h", labeller = as_labeller(vars_labels, label_parsed)) +
  geom_boxplot2() +
  geom_jitter(aes(shape = Class), alpha = .3, width = .2) +
  scale_color_manual(values = wes_palette("Royal1")) +
  scale_shape_manual(values = c(2,20)) +
  labs(title = "a") +
  theme_custom() +
  theme(axis.title.y = element_blank(), legend.position = "top",
        plot.title.position = "plot") +
  guides(color = "none")

save_plot(boxpl_pc, pdf_file = "../results/env-data/particles-boxplot.pdf", eps_file = "../results/env-data/particles-boxplot.eps", w = 16, h = 8)
saveRDS(boxpl_pc, file = "../results/env-data/particles-boxplot.rds")

# other variables
boxpl_oth <- env_epi_data %>%
  dplyr::select(class, date, weekday, aircleaner, CO2, AER, rH, Temp) %>%
  melt(c("class", "date", "weekday", "aircleaner")) %>%
  rename(`Air Cleaner` = aircleaner,
         Class = class) %>%
  ggplot(aes(x = `Air Cleaner`, color = `Air Cleaner`, y = value)) +
  facet_wrap(~ variable, scales = "free_y", nrow = 1, dir = "h", labeller = as_labeller(vars_labels, label_parsed)) +
  geom_boxplot2() +
  geom_jitter(aes(shape = Class), alpha = .3, width = .2) +
  scale_color_manual(values = wes_palette("Royal1")) +
  scale_shape_manual(values = c(2,20)) +
  theme_custom() +
  theme(axis.title.y = element_blank(), legend.position = "top") +
  guides(color = "none")
save_plot(boxpl_oth, pdf_file = "../results/env-data/otherVars-boxplot.pdf", w = 16, h = 8)
```

## Estimated change

```{r}
# log transform
env_epi_data_log <- env_epi_data %>%
  dplyr::select(-CO2, -PMtot, -Temp, -rH) %>%
  mutate(across(c(Cn, PM1, PM25, PM4, PM25, AER, n_present, cases), log))

# Bayesian linear regression model
seed0 <- 12345
draws <- list()
particle_vars <- c("Cn", "PM1", "PM25", "PM4", "PM10")
for (i in 1:length(particle_vars)) {
  f <- as.formula(paste0(particle_vars[[i]], "~ class + weekday + aircleaner + n_present + AER + cases"))
  m <- stan_glm(f, data = env_epi_data_log, seed = seed0)
  draws[[i]] <- as.data.frame(m)
}

# collect results
draws_filt <- lapply(draws, function(D) dplyr::select(D, `aircleanerYes`))
draws_filt <- do.call(rbind, draws_filt) %>%
  mutate(outcome = factor(rep(particle_vars, each = nrow(draws[[i]])), levels = rev(particle_vars))) %>%
  set_names(c("aircleaner", "outcome")) %>%
  mutate(aircleaner = (-1) * 100 * (exp(aircleaner) - 1))

# create estimation results table
effects_tbl <- draws_filt %>%
  mutate(outcome = recode(as.character(outcome), !!! vars_labels_no_units),
         outcome = factor(outcome, levels = rev(vars_labels_no_units))) %>%
  group_by(outcome) %>%
  mean_qi() %>%
  ungroup() %>%
  filter(.width == .95) %>%
  select(outcome, aircleaner, `.lower`, `.upper`) %>%
  mutate_if(is.numeric, round_k) %>%
  arrange(-row_number()) %>%
  mutate(outcome = c("CN", "PM$_{1}$", "PM$_{2.5}$", "PM$_{4}$", "PM$_{10}$")) %>%
  as.matrix() %>%
  xtable() %>%
  print.xtable(
    file = "../results/env-data/estimation-results-table.tex",
    type = "latex",
    sanitize.text.function = identity,
    include.colnames = F,
    include.rownames = F,
    only.contents = T,
    hline.after = NULL)

# create estimation results figure
effects_env_main <- draws_filt %>%
  mutate(outcome = factor(outcome, levels = (c("Cn", "PM1", "PM25", "PM4", "PM10")))) %>%
  group_by(outcome) %>%
  mean_qi(.width = c(0.5, 0.8, 0.95)) %>%
  ungroup() %>%
  arrange(desc(.width)) %>%
  mutate(`.width` = factor(`.width`, levels = rev(c(.5, .8, .95)))) %>%
  ggplot() +
  geom_rect(aes(xmin = .lower, xmax = .upper, ymin = 75, ymax = 125, fill = `.width`)) +
  geom_segment(aes(x = aircleaner, xend = aircleaner, y = 75, yend = 125), size = 1.5 / cm(1)) +
  geom_point(aes(x = aircleaner, y = 100), size = 3, fill = "white", shape = 21) +
  facet_wrap(~ outcome, nrow = 1, labeller = as_labeller(vars_labels_no_units, label_parsed)) +
  scale_fill_manual(values = wes_palette("Royal1")[c(3,4,2)]) +
  scale_y_continuous(expand = expansion(mult = c(0.25, 0.25))) +
  scale_x_continuous(limits = c(50, 100), expand = expansion(mult = c(0.1, 0.1)), breaks = c(50, 75, 100)) +
  labs(x = "Estimated reduction (%)", y = "Type", fill = "CrI", title = "b") +
  theme_custom() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(color = "white"),
        axis.ticks.y = element_blank(),
        legend.position = "top", legend.direction = "horizontal",
        legend.key.height = unit(.25, "cm"), legend.key.width = unit(.35, "cm"), 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title.position = "plot",
        panel.spacing = unit(20, units = "pt")) 
save_plot(effects_env_main, 
          pdf_file = "../results/env-data/estimation-results-figure.pdf",
          eps_file = "../results/env-data/estimation-results-figure.eps",
          w = 16, h = 5)
```