### 2023-10-17 analysis/aud_analysis.Rmd
rename flu b to ifb

---
title: "Cough Analysis"
author: "Nicolas Banholzer"
date: "2023-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(tidybayes)
library(lubridate)
library(rstanarm)
library(brms)
library(rstan)
source("../utils/epi.r")
source("../utils/bayes.r")
source("../utils/plotting.r")
source("../utils/tex.r")
```

## Data

```{r}
# load cough data 
coughs <- read.csv("../data-clean/coughs.csv", sep = ";") %>%
  mutate(class =  ifelse(class == "B", "A", "B")) %>% # correct wrong label by Simon who put Kontrollklasse 3a=A instead of Kontrollklasse 3a=B
  rename(n_coughs = `coughs.detected`) %>%
  mutate(datetime = as.POSIXct(paste(date, time), format = "%d.%m.%Y %H:%M:%S"),
         date = as.Date(datetime)) %>%
  dplyr::select(class, date, datetime, n_coughs)

# load time in the class room
time_in_room <- readRDS("../data-clean/time-in-room.rds") %>%
  dplyr::select(class, datetime, no_palas) %>%
  group_by(class) %>%
  complete(datetime = full_seq(datetime, 60)) %>%
  fill(no_palas, .direction = "down") %>%
  ungroup()

# filter time in room
coughs <- coughs %>%
  left_join(time_in_room) %>%
  filter(!no_palas)

# summarize coughs
coughs <- coughs %>%
  group_by(class, date) %>%
  summarize(n_coughs = sum(n_coughs),
            TT = n()) %>%
  ungroup()

# load epi data 
epi_data <- readRDS("../data-clean/epidemiological.rds") %>%
  dplyr::select(class, date, weekday, n_present, new_cases, aircleaner) %>%
  group_by(class) %>%
  arrange(date) %>%
  mutate(cases = cumsum(new_cases)) %>%
  ungroup() %>%
  dplyr::select(-new_cases)

# load env data
env_data <- readRDS("../data-clean/environmental.rds") %>%
  dplyr::select(class, date, AER)

# load mol data
mol_data <- readRDS("../data-clean/molecular.rds") %>%
  dplyr::select(-weekday, -aircleaner, -N, -n_class)

# merge all together
coughs <- coughs %>%
  left_join(epi_data) %>%
  left_join(env_data) 
```

## Descriptives

```{r}
# raw data
coughs %>%
  ggplot(aes(x = date, y = n_coughs / TT)) +
  geom_point(aes(color = aircleaner)) +
  geom_smooth() +
  facet_wrap(~ class)
hist(coughs$n_coughs/coughs$TT)

# global average
coughs %>%
  mutate(coughs_per_min = n_coughs / TT) %>%
  group_by(aircleaner) %>%
  summarize(mean(coughs_per_min),
            sd(coughs_per_min)) %>%
  mutate_if(is.numeric, round, 1)

# weekly coughs per min by class and condition
coughs_pl <- coughs %>%
  mutate(coughs_per_min = n_coughs / TT) %>%
  mutate(aircleaner = factor(aircleaner, levels = c("No", "Yes"))) %>%
  ggplot(aes(x = aircleaner, fill = aircleaner, y = coughs_per_min)) +
  geom_boxplot2(width.errorbar = .3, width = .4) +
  ggsignif::geom_signif(annotations = "PP(Yes<No) = 93%", y_position = 6.25, xmin = 1, xmax = 2, 
                        textsize = 8 / cm(1), vjust = -.5) +
  scale_fill_manual(values = wes_palette("Royal1")) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(add = c(0, .66))) +
  labs(title = "d", y = "Number of coughs per min", x = "Air cleaner") +
  theme_custom() +
  theme(legend.position = "none",
        plot.title.position = "plot")

coughs_pl

save_plot(coughs_pl, 
          pdf_file = "../results/cough-data/coughs-frequency-by-condition.pdf", 
          eps_file = "../results/cough-data/coughs-frequency-by-condition.eps", 
          w = 7.25, h = 6)
```

## Model

### Main

```{r}
# prepare stan data
sdl <- list()
coughs_A <- coughs %>% filter(class == "A") %>% arrange(date)
coughs_B <- coughs %>% filter(class == "B") %>% arrange(date)
sdl$Da <- nrow(coughs_A)
sdl$Db <- nrow(coughs_B)
sdl$coughsA <- coughs_A$n_coughs
sdl$coughsB <- coughs_B$n_coughs
sdl$tA <- coughs_A$TT
sdl$tB <- coughs_B$TT
sdl$airCleanerA <- ifelse(coughs_A$aircleaner == 'Yes', 1, 0)
sdl$airCleanerB <- ifelse(coughs_B$aircleaner == 'Yes', 1, 0)
sdl$aerA <- (coughs_A$AER - mean(coughs$AER)) 
sdl$aerB <- (coughs_B$AER - mean(coughs$AER))
sdl$nA <- (coughs_A$n_present - mean(coughs$n_present)) 
sdl$nB <- (coughs_B$n_present - mean(coughs$n_present)) 
sdl$casesA <- (coughs_A$cases - mean(coughs$cases)) 
sdl$casesB <- (coughs_B$cases - mean(coughs$cases)) 
sdl$weekdayA <- coughs_A %>%
  mutate(weekday = factor(weekday, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))) %>%
  dplyr::select(date, weekday) %>%
  dcast(date ~ weekday) %>%
  mutate(across(-date, ~ ifelse(is.na(.x), 0, 1))) %>%
  dplyr::select(-date, -Monday) %>%
  as.matrix()
sdl$weekdayB <- coughs_B %>%
  mutate(weekday = factor(weekday, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))) %>%
  dplyr::select(date, weekday) %>%
  dcast(date ~ weekday) %>%
  mutate(across(-date, ~ ifelse(is.na(.x), 0, 1))) %>%
  dplyr::select(-date, -Monday) %>%
  as.matrix()
sdl$s_x <- c(sd(ifelse(coughs$aircleaner == "Yes", 1, 0)), # aircleaner
             sd(ifelse(coughs$aircleaner == "Yes", 1, 0)), # class
             apply(rbind(sdl$weekdayA, sdl$weekdayB), 2, sd),
             sd(coughs$n_present),
             sd(coughs$AER),
             sd(coughs$cases))

# model
stan_file <- "../models/negbinom-coughs-model.stan"
coughsMod <- stan(stan_file, data = sdl, seed = 12345, cores = 4)

# results
raw_res <- summary(coughsMod, probs = c(0.025, 0.5, 0.975), pars = c("phi", "beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "beta6"))$summary %>%
  data.frame() %>%
  mutate_all(round, 2) 
raw_res

exp(raw_res["beta1",c("mean", "X2.5.", "X97.5.")])
exp(raw_res["beta2",c("mean", "X2.5.", "X97.5.")])

# posterior probability
coughsMod %>%
  spread_draws(beta1) %>%
  summarize(round(sum(beta1 < 0) / n() * 100))
```

### Association with molecular data

```{r}
# subset data
coughs_TueThu <- mol_data %>%
  left_join(dplyr::select(coughs, class, date, n_coughs, TT)) %>%
  left_join(epi_data) %>%
  left_join(env_data) %>%
  dplyr::select(-Neg) %>%
  left_join(coughs) %>%
  filter(!is.na(n_coughs))

# create stan data
sdl_acm <- list(
  N = nrow(coughs_TueThu),
  C = 2,
  V = 6,
  coughs = coughs_TueThu$n_coughs,
  TT = coughs_TueThu$TT,
  viruses = as.matrix(dplyr::select(coughs_TueThu, `IFB`, HRV, AdV, `CoV`, MPV, PIV))
)
sdl_acm$s_v <- mean(apply(sdl_acm$viruses, 2, sd))

# model fit
stan_file_acm <- "../models/negbinom-cough-molecular-model.stan"
coughsMolMod <- stan(stan_file_acm, data = sdl_acm, seed = 12345, cores = 4)

# results
raw_res_acm <- summary(coughsMolMod, probs = c(0.025, 0.5, 0.975), pars = c("phi", "beta0", "theta", "theta_v", "tau"))$summary %>%
  data.frame() %>%
  mutate_all(round, 2) 
raw_res_acm

# plot
viruses <- c("Average", "IFB", "HRV", "AdV", "CoV", "MPV", "PIV")
names(viruses) <- 0:6
draws_theta <- coughsMolMod %>% spread_draws(theta_v[v]) %>%
  mutate(v = recode(v, !!! viruses),
         v = factor(v, levels = viruses),
         rr = exp(theta_v))

draws_theta %>%
  dplyr::select(v, rr) %>%
  group_by(v) %>%
  mean_qi() %>%
  ungroup() %>%
  mutate_if(is.numeric, round, 2)

theta_pl <- draws_theta %>%
  ggplot(aes(y = v, x = rr, color = v)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_pointinterval(shape = 21, fill = "white") +
  scale_color_manual(values = wes_palette("IsleofDogs1")) +
  labs(x = "Relative risk") +
  theme_custom() +
  theme(axis.title.y = element_blank(), legend.position = "none",
        plot.title.position = "plot")

theta_pl

save_plot(theta_pl, 
          pdf_file = "../results/cough-data/coughs-virus-association.pdf", 
          eps_file = "../results/cough-data/coughs-virus-association.eps", 
          w = 12, h = 8)
```