### 2024-01-03 analysis/study_comaprison.Rmd
cosmetic figure changes

---
title: "Study Comparison"
author: "Nicolas Banholzer"
date: "2023-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)

library(rstan)
library(brms)
library(tidybayes)
library(modelr)

source("../utils/epi.r")
source("../utils/bayes.r")
source("../utils/plotting.r")
source("../utils/tex.r")
```

## Data

```{r}
# saliva
saliva_2022 <- readxl::read_xlsx("../data-clean/comparison/mcid-1-molecular-results.xlsx", "Saliva")
saliva_2023 <- readxl::read_xlsx("../data-clean/comparison/mcid-2-molecular-results.xlsx", "Saliva")

saliva <- rbind(saliva_2022 %>% mutate(year_num = 2022),
                saliva_2023 %>% mutate(year_num = 2023))

# bioaerosols
air_2022 <- readxl::read_xlsx("../data-clean/comparison/mcid-1-molecular-results.xlsx", "Bioaerosols")
air_2023 <- readxl::read_xlsx("../data-clean/comparison/mcid-2-molecular-results.xlsx", "Bioaerosols")

air <- rbind(air_2022 %>% mutate(year_num = 2022),
             air_2023 %>% mutate(year_num = 2023))

# air filters
# TODO: just list

# other data
odf <- readRDS("../data-clean/comparison/merged-data-both-years-all-classes.rds")
```

## Virus distribution

```{r}
saliva_sum <- saliva %>%
  mutate(IF = IFA + IFB) %>%
  dplyr::select(year_num, IF, HRV, AdV, CoV, MPV, PIV) %>%
  melt("year_num") %>%
  rename(virus = variable,
         n = value) %>%
  group_by(year_num, virus) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  filter(n > 0) %>%
  mutate(year = ifelse(year_num == 2022, "2021/22", "2022/23"),
         year = factor(year, levels = c("2021/22", "2022/23")),
         virus = factor(virus, levels = c("IF", "HRV", "AdV", "CoV", "MPV", "PIV"))) %>%
  filter(!is.na(n)) %>%
  group_by(year) %>%
  arrange(virus) %>%
  mutate(p = n / sum(n, na.rm = T),
         ymax = cumsum(p),
         ymin = lag(ymax, n = 1),
         ymin = ifelse(is.na(ymin), 0, ymin),
         labelPosition = (ymax + ymin) / 2,
         label = paste0(virus, ": ", n)) %>%
  ungroup() %>%
  mutate(labelPosition = ifelse(year == "2022/23" & virus == "MPV", labelPosition - .015, labelPosition),
         labelPosition = ifelse(year == "2021/22" & virus == "IF", ymin, labelPosition),
         labelPosition = ifelse(year == "2021/22" & virus == "AdV", ymax, labelPosition))

saliva_pl <- saliva_sum %>%
  ggplot() +
  facet_wrap(~ year, nrow = 1) +
  geom_rect(mapping = aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = virus)) +
  geomtextpath::coord_curvedpolar(theta = "y", start = 1.9*pi) +
  geom_text(mapping = aes(x = 5.25, y=labelPosition, label=label, color=virus), size = 8/cm(1)) +
  geom_text(data = data.frame(x = 1, y = 1, year = c("2021/22", "2022/23"), label = c("2021/22", "2022/23")), mapping = aes(x = x, label = label, y = y), 
            size = 8/cm(1), fontface = 2) +
  scale_x_continuous(limits = c(1, 5.25), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0),  position = "top") +
  scale_fill_manual(values = wes_palette("IsleofDogs1")) +
  scale_color_manual(values = wes_palette("IsleofDogs1")) +
  ggtitle("a") +
  theme_void() +
  theme(legend.position = "none",
        plot.title = element_text(size = 10, face = 2),
        plot.title.position = "plot",
        legend.key.height = unit(.1, "cm"),
        legend.key.width = unit(.3, "cm"),
        strip.text = element_blank()) +
  guides(fill = guide_legend(nrow = 1),
         color = "none")

saliva_pl
```

## Time distribution

```{r}
saliva_week <- saliva %>%
  mutate(IF = IFA + IFB,
         class = ifelse(year_num == 2022, ifelse(class == "D", "B", "A"), class)) %>%
  dplyr::select(year_num, week, class, intervention, IF, HRV, AdV, CoV, MPV, PIV) %>%
  melt(c("year_num", "week", "class", "intervention")) %>%
  rename(virus = variable) %>%
  group_by(year_num, class, week, intervention, virus) %>%
  summarize(n = sum(value, na.rm = T)) %>%
  ungroup()

air_week <- air %>%
  mutate(IF = IFA + IFB,
         class = ifelse(year_num == 2022, ifelse(class == "D", "B", "A"), class)) %>%
  dplyr::select(year_num, week, class, intervention, IF, HRV, AdV, CoV, MPV, PIV) %>%
  melt(c("year_num", "week", "class", "intervention")) %>%
  rename(virus = variable) %>%
  group_by(year_num, class, week, intervention, virus) %>%
  summarize(n = sum(value, na.rm = T)) %>%
  ungroup()

saliva_air_week <- rbind(saliva_week %>% mutate(type = "Saliva"),
                         air_week %>% mutate(type = "Bioaerosol")) %>%
  slice(rep(1:n(), times = n)) %>%
  mutate(year = ifelse(year_num == 2022, "Winter 2021/2022", "Winter 2022/2023"),
         year = factor(year, levels = c("Winter 2021/2022", "Winter 2022/2023")),
         virus = factor(virus, levels = c("IF", "HRV", "AdV", "CoV", "MPV", "PIV")),
         type = factor(type, levels = c("Bioaerosol", "Saliva")),
         type_num = ifelse(type == "Bioaerosol", 1, 2),
         classroom = ifelse(class == "A", "Classroom 1", "Classroom 2"),
         classroom = factor(classroom, levels = c("Classroom 1", "Classroom 2")))

backgrd_y <- saliva_air_week %>%
  group_by(year, classroom, week) %>%
  summarize(type_num = 1) %>%
  ungroup() %>%
  complete(year, classroom, week)

time_pl <- ggplot(mapping = aes(x = week, y = type_num)) +
  facet_grid(classroom ~ year) +
  geom_rect(data = backgrd_y,
            mapping = aes(xmin = week - 0.5, xmax = week + 0.5, ymin = -Inf, ymax = Inf),
            fill = ifelse(backgrd_y$week %% 2 == 0, "lightgray", "white"),
            alpha = 0.5) +
  geom_hline(data = data.frame(y = 1.5), mapping = aes(yintercept = y), linetype = "dashed") +
  geom_point(data = saliva_air_week %>%
               mutate(ps = .17) %>%
               group_by(year, classroom, week, virus, type) %>%
               mutate(type_num = type_num + cumsum(ps) - .3) %>%
               ungroup(),
             mapping = aes(color = virus),
             position = position_dodge(width = .75)) +
  scale_color_manual(values = wes_palette("IsleofDogs1")) +
  scale_x_continuous(breaks = 1:7, limits = c(0.5, 7.5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 3)) +
  annotate("text", y = 2.8, x = 0.65, label = "Saliva", hjust = 0, size = 8 / cm(1)) +
  annotate("text", y = 0.2, x = 0.65, label = "Bioaerosol", hjust = 0, size = 8 / cm(1)) +
  labs(x = "Study week", title = "b") +
  theme_bw2() +
  theme(axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = 2, size = 8),
        text = element_text(size = 8),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  guides(shape = "none",
         color = "none")

time_pl
```

## Modeling saliva-air pairs

### Checks

Look at the difference in days between positive air and saliva samples. 
* Are air samples detected later or earlier than a positive saliva samples?

```{r}
air_pos <- air %>%
  mutate(IF = IFA + IFB) %>%
  dplyr::select(-IFA, -IFB) %>%
  melt(c("year_num", "school", "class", "week", "date", "sampler", "intervention")) %>%
  filter(!is.na(value), value > 0) %>%
  mutate(class = ifelse(year_num == 2022, ifelse(class == "A/B", "A", "B"), class)) %>%
  dplyr::select(-school, -value, -intervention) %>%
  rename(date_air = date) %>%
  mutate(sample_id = 1:n())

sal_pos <- saliva %>%
  mutate(IF = IFA + IFB) %>%
  dplyr::select(-IFA, -IFB) %>%
  melt(c("year_num", "school", "class", "week", "date", "intervention")) %>%
  filter(!is.na(value), value > 0) %>%
  mutate(class = ifelse(year_num == 2022, ifelse(school == 1, "A", "B"), class)) %>%
  dplyr::select(-school, -value, -intervention) %>%
  rename(date_sal = date)

air_sal_pos <- air_pos %>% 
  dplyr::select(-week) %>%
  left_join(sal_pos %>% dplyr::select(-week)) %>%
  mutate(tdays = difftime(date_air, date_sal, units = "days")) %>%
  group_by(sample_id) %>%
  arrange(abs(tdays)) %>%
  slice(1) %>%
  ungroup()

hist(air_sal_pos$tdays %>% as.numeric, bre = 30)
```

Consider the issue of the missing VIVAS sampler in class B in 2023.
* How many weeks were there when VIVAS detected something but not Coriolis?

```{r}
air_pos_weeks <- air %>%
  mutate(IF = IFA + IFB) %>%
  dplyr::select(-IFA, -IFB) %>%
  melt(c("year_num", "school", "class", "week", "date", "sampler", "intervention")) %>%
  rename(virus = variable) %>%
  mutate(class = ifelse(year_num == 2022, ifelse(class == "A/B", "A", "B"), class)) %>%
  group_by(year_num, class, week, sampler, virus) %>%
  summarise(detected = ifelse(sum(value, na.rm = T) > 0, 1, 0)) %>%
  ungroup() %>%
  group_by(year_num, class, week, virus) %>%
  filter(any(detected == 1)) %>%
  ungroup() %>%
  filter(!(year_num == 2023 & class == "B")) %>%
  arrange(year_num, class, week, virus, sampler) %>%
  dcast(year_num + class + week + virus ~ sampler)

table(air_pos_weeks$Coriolis, air_pos_weeks$Vivas)

3 / 7
```

### Data

```{r}
# saliva bioaerosol pairs weekly
pair_df <- rbind(saliva_week %>% mutate(type = "Saliva"),
                 air_week %>% mutate(type = "Bioaerosol")) %>%
  group_by(year_num, class, virus, week, intervention, type) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  dcast(year_num + class + virus + week + intervention ~ type) %>%
  mutate(is_cov = ifelse(virus == "CoV", "yes", "no"),
         across(c(Bioaerosol, Saliva), ~ ifelse(.x > 0, "yes", "no")),
         across(c(Bioaerosol, Saliva, is_cov), ~ factor(.x, levels = c("no", "yes"))),
         year = ifelse(year_num == 2022, "Winter 2021/2022", "Winter 2022/2023"),
         year = factor(year, levels = c("Winter 2021/2022", "Winter 2022/2023")),
         virus = factor(virus, levels = c("IF", "HRV", "AdV", "CoV", "MPV", "PIV")),
         intervention = factor(intervention, levels = c("none", "masks", "air cleaners")))

# add ventilation
vent <- left_join(
  air %>% 
    mutate(date = as.Date(date),
           class = ifelse(class == "A/B", "A", ifelse(class == "D", "B", class))) %>%
    group_by(year_num, class, week, date) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(year_num, class, week, date),
  odf %>%
    group_by(school, classroom, date) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(year_num = lubridate::year(date),
           class = ifelse(classroom == "AB", "A",
                          ifelse(year_num == 2022 & school == "School 2" & classroom == "A", "B", classroom))) %>%
    dplyr::select(year_num, class, date, co2.max) 
) %>%
  group_by(year_num, class, week) %>%
  summarize(maxCO2_imputed = ifelse(is.na(co2.max), T, F), 
            co2.max = mean(co2.max, na.rm = T)) %>%
  ungroup() %>%
  group_by(year_num) %>%
  mutate(co2.max = ifelse(is.na(co2.max), mean(co2.max, na.rm = T), co2.max)) %>%
  ungroup() %>%
  rename(maxCO2 = co2.max)

pair_df <- left_join(pair_df, vent)

pair_df %>%
  group_by(year_num, class, week) %>%
  slice(1) %>%
  ungroup() %>%
  filter(!maxCO2_imputed) %>%
  group_by(year_num) %>%
  summarize(mean(maxCO2), sd(maxCO2)) %>%
  mutate_if(is.numeric, round, 0)

# create stan data file
stanDF <- list(
  N = nrow(pair_df),
  K_virus = n_distinct(pair_df$virus),
  bioaerosol = ifelse(pair_df$Bioaerosol == "yes", 1, 0),
  saliva = ifelse(pair_df$Saliva == "yes", 1, 0),
  cov = ifelse(pair_df$is_cov == "yes", 1, 0),
  virus = as.integer(pair_df$virus),
  masks = ifelse(pair_df$intervention == "masks", 1, 0),
  air_cleaners = ifelse(pair_df$intervention == "air cleaners", 1, 0),
  co2max = (pair_df$maxCO2 - mean(pair_df$maxCO2)) / sd(pair_df$maxCO2) 
)
```

### Priors

```{r}
# add prior scales
stanDF$s_saliva <- sd(stanDF$bioaerosol) / sd(stanDF$saliva)
stanDF$s_cov <- sd(stanDF$bioaerosol) / sd(stanDF$cov)
stanDF$s_masks <- .5 * sd(stanDF$bioaerosol) / sd(stanDF$masks)
stanDF$s_air <- .5 * sd(stanDF$bioaerosol) / sd(stanDF$air_cleaners)
stanDF$s_co2 <- sd(stanDF$bioaerosol)

# add prior skew
stanDF$t_masks <- .75
stanDF$t_air <- .75
stanDF$t_co2 <- .25

# add offset
stanDF$offsets = ifelse(pair_df$year_num == 2022, ifelse(pair_df$class == "A", log(2), log(1)), 
                        ifelse(pair_df$class == "A", log(1), log(.5)))

# informative priors
#' positive effect of saliva based on conditional frequencies

table(pair_df$Saliva, pair_df$Bioaerosol) 
# 5/3 odds of positive bioaerosol for positive saliva
stanDF$m_saliva <- sum(pair_df$Saliva[pair_df$Bioaerosol == "yes"] == "yes") / sum(pair_df$Saliva[pair_df$Bioaerosol == "yes"] == "no")

#' 80% prior prob. for a negative effect of interventions
#' smaller effects sizes expected for interventions

stanDF$m_masks <- optim(par = 0, 
                 fn = function(l, s, k, p) {
                   q <- qasym_laplace(p, l, s, k)
                   abs(q)
                   }, 
                 s = stanDF$s_masks, k = stanDF$t_masks, p = 0.9)$par[1] 
stanDF$m_air <- optim(par = 0, 
                 fn = function(l, s, k, p) {
                   q <- qasym_laplace(p, l, s, k)
                   abs(q)
                   }, 
                 s = stanDF$s_air, k = stanDF$t_air, p = 0.9)$par[1]

#' same prior for co2 as for intervention effects, but positive

stanDF$m_co2 <- optim(par = 0, 
                 fn = function(l, s, k, p) {
                   q <- qasym_laplace(p, l, s, k)
                   abs(q)
                   }, 
                 s = stanDF$s_co2, k = stanDF$t_co2, p = 0.1)$par[1]

#' uninformative prior for mean of cov
stanDF$m_cov <- 1
```

```{r}
a  <- seq(-5, 5, .1)
plot(a, dasym_laplace(a, stanDF$m_masks, stanDF$s_masks, stanDF$t_masks))
```

### Model

```{r}
# fit model
mod <- stan(file = "../models/comparison/air-detection-model.stan",
            data = stanDF,
            cores = 4, chains = 4, iter = 2000,
            seed = 12345)

summary(mod, pars = c("alpha", 
                      # "tau_virus", paste0("beta_virus[", 1:stanDF$K_virus, "]"), 
                      "beta_saliva", 
                      "beta_cov", 
                      "beta_masks", 
                      "beta_air_cleaners", "beta_co2max"))$summary %>%
  data.frame() %>%
  add_rownames() %>%
  as_data_frame() %>%
  mutate_if(is.numeric, round, 2)
```

### Results

```{r}
# adjusted OR of mask wearing
mod %>%
  spread_draws(beta_cov) %>%
  mutate(beta_cov = exp(beta_cov)) %>%
  median_qi() %>%
  mutate_if(is.numeric, round, 1)

pred_prob <- mod %>%
  spread_draws(logit_cov, logit_nonCov) %>%
  mutate(CoV = invlogit(logit_cov),
         `non-Cov` = invlogit(logit_nonCov)) %>%
  dplyr::select(-logit_cov, -logit_nonCov) %>%
  melt(c(".chain", ".iteration", ".draw"))

prob_pl <- ggplot(mapping = aes(x = variable, y = value, color = variable)) +
  geom_errorbar(data = pred_prob %>% group_by(variable) %>% median_qi(value),
                mapping = aes(ymin = .lower, ymax = .upper, color = variable),
                width = .25) +
  geom_boxplot(data = pred_prob,
               position = position_dodge(width = .5), outlier.shape = NA, coef = 0, width = 0.25) +
  stat_summary(data = pred_prob %>% dplyr::select(value, variable),
               mapping = aes(x = variable, y = value, color = variable), geom = "point", fun = "median", 
               position = position_dodge2(width = .5), size = 2, shape = 23, fill = "white") +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), labels = function(x) x * 100) +
  scale_color_manual(values = c(wes_palette("IsleofDogs1")[4], "darkred")) +
  #scale_fill_manual(values = c(wes_palette("IsleofDogs1")[4], "darkred")) +
  labs(y = "Prob. of airborne detection (%)", title = "c") +
  theme_custom() +
  theme(legend.position = "none",
        axis.title.x = element_blank())

prob_pl

# CoV vs non-CoV
pred_prob %>%
  group_by(variable) %>%
  median_qi() %>%
  mutate_if(is.numeric, function(x) round(x * 100) )
```

```{r}
design <- "12
33"
comparison <- saliva_pl + prob_pl + time_pl + plot_layout(design = design, widths = c(11, 5), heights = c(6, 9))
ggsave(plot = comparison, file = "../results/comparison.pdf", width = 16 / cm(1), height = 15 / cm(1))
```

<!-- ## No. of positive samples -->

<!-- ```{r} -->
<!-- pos_samples <- data.frame( -->
<!--   year_num = factor(c(rep(2022, 3), rep(2023, 3)), levels = c(2022, 2023)), -->
<!--   type = factor(rep(c("Saliva", "Bioaerosols", "HEPA-filter"), times = 2), levels = c("Saliva", "Bioaerosols", "HEPA-filter")), -->
<!--   n_pos = c(c(21, 10, 8), c(50, 2, 4)), -->
<!--   n_tot = c(c(262, 130, 40), c(448, 105, 160)) -->
<!-- ) %>% -->
<!--   mutate(n_neg = n_tot - n_pos, -->
<!--          year = ifelse(year_num == 2022, "Winter 2021/2022", "Winter 2022/2023"), -->
<!--          year = factor(year, levels = c("Winter 2021/2022", "Winter 2022/2023"))) %>% -->
<!--   dplyr::select(-n_tot,-n_neg) %>% -->
<!--   reshape2::melt(c("year_num", "year", "type")) -->

<!-- pos_samples_pl <- pos_samples %>% -->
<!--   ggplot(aes(y = value, x = year_num, fill = year)) + -->
<!--   geom_bar(stat = "identity") + -->
<!--   geom_text(aes(label = value, color = year), size = 8 / cm(1), vjust = -1) + -->
<!--   facet_wrap(~ type, nrow = 1) + -->
<!--   scale_y_continuous(expand = expansion(add = c(0,5))) + -->
<!--   scale_fill_manual(values = RColorBrewer::brewer.pal(9, "PuBu")[c(4,9)]) + -->
<!--   scale_color_manual(values = RColorBrewer::brewer.pal(9, "PuBu")[c(4,9)]) + -->
<!--   labs(y = "Frequency of positive samples", title = "b") + -->
<!--   theme_custom() + -->
<!--   theme(legend.position = "none", -->
<!--         plot.title.position = "plot", -->
<!--         strip.text = element_text(face = 2, vjust = 2, margin = margin(t = 10, b = 10)), -->
<!--         panel.spacing.x = unit(1, "cm"), -->
<!--         axis.title.x = element_blank(), -->
<!--         panel.grid.major = element_blank(), -->
<!--         panel.grid.minor = element_blank()) -->

<!-- pos_samples_pl -->
<!-- ``` -->

<!-- ```{r} -->
<!-- comparison <- grid.arrange(saliva_pl, pos_samples_pl, nrow = 2, heights = c(6, 8), widths = 11) -->
<!-- ggsave(plot = comparison, file = "../results/comparison.pdf", width = 11.5 / cm(1), height = 14 / cm(1)) -->
<!-- ``` -->

<!-- ## Absences ## -->

<!-- ```{r} -->
<!-- mcid_2 <- readRDS("../data-clean/redcap-long.rds") %>% -->
<!--   filter(respiratory_infection == "yes") %>% -->
<!--   mutate(absent_days = date_back - date_absent) -->

<!-- # 35 absences related to respiratory infection among 38 students -->

<!-- summary(as.numeric(mcid_2$absent_days)) -->

<!-- # on average 3.0 days out of school per absence (but weekends should be kept in mind) -->

<!-- mcid_1 <- read_csv("../data-clean/mcid-1-redcap-cases.csv") %>% -->
<!--   filter(!is_teacher, -->
<!--          school_class == "School 1 (A)" | school_class == "School 2 (D)") %>% -->
<!--   mutate(absent_days = date_end - date_start) -->

<!-- # 40 absences related to respiratory infection among 44 students -->

<!-- summary(as.numeric(mcid_1$absent_days)) -->

<!-- #' on average 4.1 days out of school per absence (but weekends should be kept in mind) -->
<!-- #' there is also an outlier with 30 days, but the median is still higher  -->

<!-- rbind(mcid_1 %>% mutate(year = 2022) %>% dplyr::select(year, absent_days), -->
<!--       mcid_2 %>% mutate(year = 2023) %>% dplyr::select(year, absent_days)) %>% -->
<!--   mutate(absent_days = as.numeric(absent_days)) %>% -->
<!--   ggplot(aes(fill = factor(year), x = year, y = absent_days)) + -->
<!--   geom_boxplot() + -->
<!--   scale_y_sqrt(breaks = c(0, 1:7, 10, 20, 30), limits = c(1,30)) -->
<!-- ``` -->

<!-- ## Particle reduction -->

<!-- ```{r} -->
<!-- vars <- c("CO2", "rH", "Temp", "AER", "Cn", "PM1", "PM25", "PM4", "PM10") -->
<!-- vars_labels_no_units <- c("CO[2]", "Relative~humidity", "Temperature", "Air~exchange~rate", -->
<!--                           "CN", "PM[1]", "PM[2.5]", "PM[4]", "PM[10]") -->
<!-- names(vars_labels_no_units) <- vars -->

<!-- particle <- data.frame( -->
<!--   year = factor(c(rep(2022, 15), rep(2023, 15)), levels = c(2022, 2023)), -->
<!--   variable = factor(c(rep(c("Cn", "PM1", "PM25", "PM4", "PM10"), each = 3, times = 2)), levels = c("Cn", "PM1", "PM25", "PM4", "PM10")), -->
<!--   interval = rep(c("Mean", "Lower", "Upper"), times = 5), -->
<!--   estimate = c(c(39.24, 3.88, 68.80,  -->
<!--                  41.05, 6.62, 67.98, -->
<!--                  45.12, 19.66, 63.16, -->
<!--                  46.49, 27.26, 61.72, -->
<!--                  46.60, 29.99, 59.49), -->
<!--                c(76.83, 63.20, 86.28, -->
<!--                  76.79, 64.84, 85.46, -->
<!--                  72.86, 64.24, 79.66, -->
<!--                  70.77, 63.46, 76.97, -->
<!--                  99.99, 99.94, 100.00)) -->
<!-- ) %>% -->
<!--   dcast(year + variable ~ interval) %>% -->
<!--   ggplot(aes(y = year, color = year)) + -->
<!--   facet_wrap(~ variable, nrow = 1, dir = "h", labeller = as_labeller(vars_labels_no_units, label_parsed)) + -->
<!--   geom_errorbar(mapping = aes(xmin = Lower, xmax = Upper), lineend = "square", size = 1, width = .05) + -->
<!--   geom_point(mapping = aes(x = Mean), shape = 23, size = 2, fill = "white") + -->
<!--   scale_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) + -->
<!--   scale_x_continuous(expand = expansion(mult = c(0, 0.1)), breaks = c(0, 50, 100), limits = c(0, 100)) + -->
<!--   labs(y = "Study year", x = "Estimated reduction in particle concentration (%)", title = "c") +  -->
<!--   theme_custom() + -->
<!--   theme(legend.position = "none", -->
<!--         plot.title.position = "plot", -->
<!--         panel.spacing = unit(.5, "lines")) -->

<!-- particle -->
<!-- ``` -->

<!-- ## Risk ratio -->

<!-- ```{r} -->
<!-- rr <- data.frame( -->
<!--   year = factor(c(rep(2022, 3), rep(2023, 3)), levels = c(2022, 2023)), -->
<!--   interval = rep(c("Mean", "Lower", "Upper"), times = 2), -->
<!--   estimate = c(c(1.00, 0.22, 2.067), c(0.72, 0.44, 1.15)) -->
<!-- ) %>% -->
<!--   dcast(year ~ interval) %>% -->
<!--   ggplot(aes(y = year, color = year)) + -->
<!--   geom_vline(aes(xintercept = 1), linetype = "dashed") + -->
<!--   geom_errorbar(mapping = aes(xmin = Lower, xmax = Upper), lineend = "square", size = 1, width = .05) + -->
<!--   geom_point(mapping = aes(x = Mean), shape = 23, size = 2, fill = "white") + -->
<!--   labs(y = "Study year", x = "ARR of infection with air cleaners (%)", title = "d") +  -->
<!--   scale_color_manual(values = wes_palette("Darjeeling2")[c(2, 5)]) + -->
<!--   theme_custom() + -->
<!--   theme(legend.position = "none", -->
<!--         plot.title.position = "plot") -->

<!-- rr   -->
<!-- ``` -->

<!-- ## Combined  -->

<!-- ```{r} -->
<!-- comparison <- grid.arrange(saliva, pos_samples, particle, rr, ncol = 2, nrow = 2, heights = c(6, 4), widths = c(10,6)) -->
<!-- ggsave(plot = comparison, file = "../results/comparison.pdf", width = 16 / cm(1), height = 10 / cm(1)) -->
<!-- ``` -->